Convolutional Neural Networks and Volcano Plots: Screening and Prediction of Two-Dimensional Single-Atom Catalysts for CO$_2$ Reduction Reactions

### 2024-01-05 paper/main_sections/4.2_volcano_plots.tex
add cross-ref in main text

Catalyst activity volcano plots, built upon the scaling relations proposed by Peterson and Nørskov's \cite{peterson2012activity},
goes beyond assessment of candidatures' energetics only but to provide informative guidance on discovering catalysts candidates with superior performance \cite{balandin1969modern, deutschmann2000heterogeneous}.
Our endeavor to improve the accuracy of volcano plots leads to the development of the hybrid-descriptor,
which encompasses both C- and O- bonding types of CO$_2$RR intermediates within the scaling relations.
This adaptation arose from the intuition that most CO$_2$RR intermediates are featured with both bonding types rather than any single one in dominance.
By integrating both bonding featured descriptors into the hybrid approach, we found the coefficient of determination ($R^2$) was increased by 0.1387.
Detailed information about the scaling parameters and performance enhancements are provided in Supplementary Tables 13 and Supplementary Figure 10.
This improvement to some extent underscored our premise that CO$_2$RR intermediates typically exhibit chemical similarities encompassing both O- and C-type descriptors.
After taking this dual nature into consideration, our hybrid-descriptor method can evaluate the intermediates adsorption energies more accurately.
Importantly, the introduction of hybrid descriptors is computationally affordable compared to DFT methods,
as it involves only a set of linear regressions to identify the optimal mixing ratio, as elaborated in the Methods section.

Expanding upon the revised scaling relations, we introduce an activity volcano plot and a rate-determining step (RDS) volcano plot for the CO$_2$RR process, shown in Figure 3a and Supplementary Figure 11.
Given that HER is a key competing reaction with CO$_2$RR \cite{goyal2020competition}, we also incorporated a selectivity volcano plot, presented in Supplementary Figure 12, to offer a comprehensive analysis of catalyst candidates.
In adherence to Peterson and Nørskov's original framework \cite{peterson2012activity}, our choice of x and y axes descriptors aligned with $\textit{G}_{\text{ads} \ast \text{CO}}$ and $\textit{G}_{\text{ads} \ast \text{OH}}$.
Further elaboration on how the limiting potentials were evaluated across the volcano plot can be found in the Methods section.
Among the candidatures explored, Ge@g-C$_3$N$_4$ emerged as the most promising catalyst, demonstrating theoretical activity at a limiting potential of -1.2126 eV.
Notably, the observed selectivity trend paralleled the activity trend, suggesting that catalysts exhibiting high theoretical activity are also resistant to HER side reactions.
The RDS volcano plot (Supplementary Figure 11) indicated protonation of *CO as the rate-determining step for nearly optimal candidates.
Moreover, the activity volcano plot predicted an optimal catalyst featuring a theoretical limiting potential of -1.0516 V, at $\textit{G}_{\text{ads} \ast \text{CO}}$ of 0.0589 eV and $\textit{G}_{\text{ads} \ast \text{OH}}$ of -4.0075 eV, positioned at the right-center of the volcano plot.
Results from RDS volcano plot indicate that protonation of *CO likely serves as the RDS for CO$_2$RR.
Consequently, it should be the focal point for future catalyst designs.
The activity and selectivity volcano plots imply that enhancing theoretical activity could be achieved by decreasing CO affinity while increasing OH affinity with the catalysts.
This result suggests a perspective: weak binding of CO to the catalyst surface might facilitate the rotation of C-O backbone from the “upright” configuration in *CO to the “horizontal” orientation in *CHO \cite{peterson2010copper},
and therefore facilitating dynamics and protonation of the *CO species.

@@

Additionally, we mapped the theoretical limiting potentials of the analyzed catalyst candidates onto periodic tables, shown in Figure 3b and Supplementary Figure 13 to 17.
Our analyses do not reveal consistent trends within groups or periods, nor did they exhibit evident patterns across substrates.
These observations suggest intricate interactions between supported metal atoms and substrates.
The absence of straightforward trends underscores the need for nuanced explorations in understanding these interactions like the CNN model.

>>>>

Catalyst activity volcano plots, built upon the scaling relations proposed by Peterson and Nørskov's \cite{peterson2012activity},
goes beyond assessment of candidatures' energetics only but to provide informative guidance on discovering catalysts candidates with superior performance \cite{balandin1969modern, deutschmann2000heterogeneous}.
Our endeavor to improve the accuracy of volcano plots leads to the development of the hybrid-descriptor,
which encompasses both C- and O- bonding types of CO$_2$RR intermediates within the scaling relations.
This adaptation arose from the intuition that most CO$_2$RR intermediates are featured with both bonding types rather than any single one in dominance.
By integrating both bonding featured descriptors into the hybrid approach, we found the coefficient of determination ($R^2$) was increased by 0.1387.
Detailed information about the scaling parameters and performance enhancements are provided in Supplementary Tables 13 and Supplementary Figure 10 (TODO:).
This improvement to some extent underscored our premise that CO$_2$RR intermediates typically exhibit chemical similarities encompassing both O- and C-type descriptors.
After taking this dual nature into consideration, our hybrid-descriptor method can evaluate the intermediates adsorption energies more accurately.
Importantly, the introduction of hybrid descriptors is computationally affordable compared to DFT methods,
as it involves only a set of linear regressions to identify the optimal mixing ratio, as elaborated in the Methods section.

Expanding upon the revised scaling relations, we introduce an activity volcano plot and a rate-determining step (RDS) volcano plot for the CO$_2$RR process, shown in \cref{main_fig3:volcano}a and Supplementary Figure 11 (TODO:).
Given that HER is a key competing reaction with CO$_2$RR \cite{goyal2020competition}, we also incorporated a selectivity volcano plot, presented in Supplementary Figure 12 (TODO:), to offer a comprehensive analysis of catalyst candidates.
In adherence to Peterson and Nørskov's original framework \cite{peterson2012activity}, our choice of x and y axes descriptors aligned with $\textit{G}_{\text{ads} \ast \text{CO}}$ and $\textit{G}_{\text{ads} \ast \text{OH}}$.
Further elaboration on how the limiting potentials were evaluated across the volcano plot can be found in the Methods section.
Among the candidatures explored, Ge@g-C$_3$N$_4$ emerged as the most promising catalyst, demonstrating theoretical activity at a limiting potential of -1.2126 eV.
Notably, the observed selectivity trend paralleled the activity trend, suggesting that catalysts exhibiting high theoretical activity are also resistant to HER side reactions.
The RDS volcano plot (Supplementary Figure 11 TODO:) indicated protonation of *CO as the rate-determining step for nearly optimal candidates.
Moreover, the activity volcano plot predicted an optimal catalyst featuring a theoretical limiting potential of -1.0516 V, at $\textit{G}_{\text{ads} \ast \text{CO}}$ of 0.0589 eV and $\textit{G}_{\text{ads} \ast \text{OH}}$ of -4.0075 eV, positioned at the right-center of the volcano plot.
Results from RDS volcano plot indicate that protonation of *CO likely serves as the RDS for CO$_2$RR.
Consequently, it should be the focal point for future catalyst designs.
The activity and selectivity volcano plots imply that enhancing theoretical activity could be achieved by decreasing CO affinity while increasing OH affinity with the catalysts.
This result suggests a perspective: weak binding of CO to the catalyst surface might facilitate the rotation of C-O backbone from the “upright” configuration in *CO to the “horizontal” orientation in *CHO \cite{peterson2010copper},
and therefore facilitating dynamics and protonation of the *CO species.

@@

Additionally, we mapped the theoretical limiting potentials of the analyzed catalyst candidates onto periodic tables, shown in \cref{main_fig3:volcano}b and Supplementary Figure 13 to 17 (TODO:).
Our analyses do not reveal consistent trends within groups or periods, nor did they exhibit evident patterns across substrates.
These observations suggest intricate interactions between supported metal atoms and substrates.
The absence of straightforward trends underscores the need for nuanced explorations in understanding these interactions like the CNN model.

### 2024-01-05 paper/main_sections/4.3_validation.tex
add cross-ref in main text

Historically, interpretability has posed challenges for deep learning methods, often rendering their inner workings akin to black boxes \cite{zhang2018interpretable, zhang2018visual, savage2022breaking}.
The occlusion technique, widely utilized in Computer Vision \cite{zeiler2014visualizing, kortylewski2020combining, wang2020robust}, involves systematically masking sections of the input matrix to assess their impact on the final prediction score.
In our work, we introduce an orbitalwise occlusion method to discern individual orbital contribution to adsorption energy within the CNN model, drawing inspiration from Fung's pioneer integration of this technique into eDOS analysis \cite{fung2021machine}.
In this work, occlusion experiments involve orbitalwise masking specific eDOS, which are then fed into the CNN model to observe resulting disturbance in adsorption energy.
A visual representation of this process can be found in Supplementary Figure 21. Conceptually, this experiment simulates the effect of shielding electronic states, allowing investigation of the potential contribution to adsorption energy from these states.
Additionally, we attempted to extract the planewave coefficients of Kohn-Sham orbital for wavefunction visualization.
This validation step was undertaken to confirm the spatial distribution of the orbitals pinpointed by the occlusion experiments.

To validate our method, we conducted occlusion experiments on Cr@g-C$_3$N$_4$, Co@g-C$_3$N$_4$ and Cu@g-C$_3$N$_4$.
Each of these supported metals possesses 3d subshells with different electronic configurations: half-filled, partially filled and fully filled, allowing us to test the universality of our approach.
The results of the occlusion experiment on the supported Cr single atom in Cr@g-C$_3$N$_4$ are given in Figure 4a.
The $d_{x^2-y^2}$ orbital emerges as dominant in CO interaction.
Shielding electrons centered around -2 eV significantly alter the adsorption energy, indicating the potency of this orbital.
The real-space wavefunction visualization in Figure 4d corroborates these findings, highlighting the interaction between the $d_{x^2-y^2}$ orbital of single metal atom and the adsorbate.
This suggests the $d_{x^2-y^2}$ orbital of supported Co predominantly influence the interaction with CO, and thus shielding electrons from this orbital markedly disturb the adsorption energy.
Additionally, occlusion experiments were performed on the Co $d_xy$ orbital of Co@g-C$_3$N$_4$ (Figure 4b) and the $Cu d_xy$ orbital of Cu@g-C$_3$N$_4$ (Figure 4c).
In both cases, the $d_xy$ orbital was identified as the dominant contributor.
While identifying the dominant orbital is crucial, these findings also suggest that interactions with adsorbates occurring out-of-plane are influenced by in-plane orbitals as well.
Importantly, it becomes evident that d orbitals do not universally impact interactions with adsorbates, emphasizing the need for case-specific discussions.

To validate the chemical relevance of predictions made by CNN model, we conducted COHP analysis, a theoretical bond-detecting tool used to identify bonding and anti-bonding contributions to band-structure energy \cite{deringer2011crystal}.
In this research, interactions between the supported metal atom and atoms within a 5 \text{\AA} radius were calculated to understand its bonding environment, as detailed in the Methods section.
The results of the occlusion experiment on the Co $d_xy$ orbital of Co@g-C$_3$N$_4$, identified as the dominant orbital via occlusion experiments, were compared with the COHP analysis, as shown in Figure 4g and Figure 4i.
The occlusion experiment pinpointed strong metal-adsorbate interactions at energy levels of -1 eV and -6.3 eV, which were confirmed by the COHP analysis.
We then investigated the non-dominant Co $d_xz$ orbital, and the COHP analysis highlighted interactions at -1 eV and -6.5 eV, aligning with the findings from the occlusion experiments.

>>>>

Historically, interpretability has posed challenges for deep learning methods, often rendering their inner workings akin to black boxes \cite{zhang2018interpretable, zhang2018visual, savage2022breaking}.
The occlusion technique, widely utilized in Computer Vision \cite{zeiler2014visualizing, kortylewski2020combining, wang2020robust}, involves systematically masking sections of the input matrix to assess their impact on the final prediction score.
In our work, we introduce an orbitalwise occlusion method to discern individual orbital contribution to adsorption energy within the CNN model, drawing inspiration from Fung's pioneer integration of this technique into eDOS analysis \cite{fung2021machine}.
In this work, occlusion experiments involve orbitalwise masking specific eDOS, which are then fed into the CNN model to observe resulting disturbance in adsorption energy.
A visual representation of this process can be found in Supplementary Figure 21 (TODO:). Conceptually, this experiment simulates the effect of shielding electronic states, allowing investigation of the potential contribution to adsorption energy from these states.
Additionally, we attempted to extract the planewave coefficients of Kohn-Sham orbital for wavefunction visualization.
This validation step was undertaken to confirm the spatial distribution of the orbitals pinpointed by the occlusion experiments.

To validate our method, we conducted occlusion experiments on Cr@g-C$_3$N$_4$, Co@g-C$_3$N$_4$ and Cu@g-C$_3$N$_4$.
Each of these supported metals possesses 3d subshells with different electronic configurations: half-filled, partially filled and fully filled, allowing us to test the universality of our approach.
The results of the occlusion experiment on the supported Cr single atom in Cr@g-C$_3$N$_4$ are given in \cref{main_fig4:validation}a.
The $d_{x^2-y^2}$ orbital emerges as dominant in CO interaction.
Shielding electrons centered around -2 eV significantly alter the adsorption energy, indicating the potency of this orbital.
The real-space wavefunction visualization in \cref{main_fig4:validation}d corroborates these findings, highlighting the interaction between the $d_{x^2-y^2}$ orbital of single metal atom and the adsorbate.
This suggests the $d_{x^2-y^2}$ orbital of supported Co predominantly influence the interaction with CO, and thus shielding electrons from this orbital markedly disturb the adsorption energy.
Additionally, occlusion experiments were performed on the Co $d_xy$ orbital of Co@g-C$_3$N$_4$ (\cref{main_fig4:validation}b) and the $Cu d_xy$ orbital of Cu@g-C$_3$N$_4$ (\cref{main_fig4:validation}c).
In both cases, the $d_xy$ orbital was identified as the dominant contributor.
While identifying the dominant orbital is crucial, these findings also suggest that interactions with adsorbates occurring out-of-plane are influenced by in-plane orbitals as well.
Importantly, it becomes evident that d orbitals do not universally impact interactions with adsorbates, emphasizing the need for case-specific discussions.

To validate the chemical relevance of predictions made by CNN model, we conducted COHP analysis, a theoretical bond-detecting tool used to identify bonding and anti-bonding contributions to band-structure energy \cite{deringer2011crystal}.
In this research, interactions between the supported metal atom and atoms within a 5 \text{\AA} radius were calculated to understand its bonding environment, as detailed in the Methods section.
The results of the occlusion experiment on the Co $d_xy$ orbital of Co@g-C$_3$N$_4$, identified as the dominant orbital via occlusion experiments, were compared with the COHP analysis, as shown in \cref{main_fig4:validation}g and \cref{main_fig4:validation}i.
The occlusion experiment pinpointed strong metal-adsorbate interactions at energy levels of -1 eV and -6.3 eV, which were confirmed by the COHP analysis.
We then investigated the non-dominant Co $d_xz$ orbital, and the COHP analysis highlighted interactions at -1 eV and -6.5 eV, aligning with the findings from the occlusion experiments.

### 2024-01-05 paper/main_sections/4.4_shifting_exp.tex
add cross-ref in main text

Figure 5 illustrates that supported Ga and Ge, the sole p-block metals in our study, exhibited no response to d orbital shifting.
This observation is consistent with their lack of d-electrons in their valence shells, reinforcing the reliability of our shifting experiments.
We also investigated potential disturbance from p orbital shifting, as shown in Supplementary Figure 31 and found no significant response.
This implies that Ge@g-C$_3$N$_4$ holds promise as a CO$_2$RR electrocatalyst, given its high activity and resilience to electronic structure disturbances.
Nevertheless, fine-tuning its performance through eDOS modulation might pose a challenge.
In contrast, supported Fe, Cu and Zn display strong sensitivity to eDOS shifting.
Interestingly, the adsorption energy consistently shifted towards more positive values, indicating a weakening of the catalyst-adsorbate interaction, regardless of the shifting direction within the investigated energy range.
This underscores the intricate interplay between the supported metal atom and the substrate, emphasizing the necessity of a tool like the CNN model to predict such disturbances.
Supported Cr exhibited a contrary response to eDOS shifting, strengthening the interaction with the CO adsorbate regardless of the shifting direction.
Supported Ti, V, Co and Ni showed weak response to eDOS shifting, suggesting the need for alternative methods beyond manipulating their eDOS to regulate their interactions with CO adsorbate.

>>>>

\cref{main_fig5:shifting} illustrates that supported Ga and Ge, the sole p-block metals in our study, exhibited no response to d orbital shifting.
This observation is consistent with their lack of d-electrons in their valence shells, reinforcing the reliability of our shifting experiments.
We also investigated potential disturbance from p orbital shifting, as shown in Supplementary Figure 31 (TODO:) and found no significant response.
This implies that Ge@g-C$_3$N$_4$ holds promise as a CO$_2$RR electrocatalyst, given its high activity and resilience to electronic structure disturbances.
Nevertheless, fine-tuning its performance through eDOS modulation might pose a challenge.
In contrast, supported Fe, Cu and Zn display strong sensitivity to eDOS shifting.
Interestingly, the adsorption energy consistently shifted towards more positive values, indicating a weakening of the catalyst-adsorbate interaction, regardless of the shifting direction within the investigated energy range.
This underscores the intricate interplay between the supported metal atom and the substrate, emphasizing the necessity of a tool like the CNN model to predict such disturbances.
Supported Cr exhibited a contrary response to eDOS shifting, strengthening the interaction with the CO adsorbate regardless of the shifting direction.
Supported Ti, V, Co and Ni showed weak response to eDOS shifting, suggesting the need for alternative methods beyond manipulating their eDOS to regulate their interactions with CO adsorbate.

### 2024-01-05 paper/main_sections/5_methods.tex
add cross-ref in main text

\subsection{Catalyst models}
Supported SACs were modelled by anchoring individual metal atoms onto various 2D substrates, including graphitic carbon nitride (g-C$_3$N$_4$), nitrogen-doped graphene, graphene with dual-vacancy, black phosphorous, boron nitride, and monolayer C$_2$N.
The specific single metal atoms studied are outlined in Supplementary Figure 1.
To prevent interactions between adjacent images along the z-axis, vacuum layers of 15 \text{\AA} were introduced.

@@

\subsection{Hybrid descriptor for volcano plots}
Hybrid descriptors incorporate both descriptors within a single equation, enhancing the accuracy of scaling relationship with minimal computational overhead, as demonstrated in Supplementary Figure 10 and Supplementary Table 13.
Comprehensive analyses are presented Supplementary Section 1.5. This leads to the formulation of the scaling relation:
\begin{equation*}
G_{\text{ads}} Z = k[\theta^* G_{\text{ads}} \ce{CO} + (1-\theta^*) G_{\text{ads}} \ce{OH}] + c_Z
\end{equation*}
Here, $\theta$ represents the mixing ratio of the CO descriptor, $\textit{Z}$ denotes any adsorbate, and $\textit{k}$ and $\textit{c}_{Z}$ are adsorbate-specific scaling parameters.
The optimal mixing ratio $\theta$ is determined through iterative exploration across the range, utilizing a step length of 0.01.

@@

\subsection{CNN architecture and hyperparameter tuning}
We designed our CNN architecture drawing inspiration from its successful implementation in tasks involving electrocardiogram signal classification \cite{weimann2021transfer}, as illustrated in Figure 2b and 2c.
This multi-branched CNN comprises nine branches tailored to handle eDOS orbitals while ensuring a reduced memory footprint.

@@

\subsection{Occlusion experiments}
Supplementary Figure 21 visually illustrates the occlusion experiment methodology.
To maintain consistent shapes, zero-padding was applied to the input eDOS arrays.
Focusing on understanding the influence of metals on adsorption behavior, a masker with dimensions [width, 1, 1] was employed along the metal channel to eliminate cross-orbital interactions.
Here, "width" denotes the masker width, and detailed explanations for determining the width are provided in Supplementary Section 2.6.
Through recursive application of these maskers, variations in predicted adsorption energies from the CNN model were recorded, generating an occlusion array.

\subsection{Shifting experiments}
Supplementary Figure 30 illustrates the implemented shifting experiment protocol.
In this study, the initial electronic density of states underwent controlled shifting along the energy axis.
This protocol ranges from -1 eV to 1 eV, with a step length of 0.005 eV per image.
To maintain a uniform energy window, each image underwent cropping and padding procedures.
The resulted output arrays took the form of [400, 1] arrays for simultaneous orbital shifting and [400, 9] arrays for individual orbital shifting.
The energy range for shifting was chosen to induce a moderate disturbance to adsorption energy.

>>>>

\subsection{Catalyst models}
Supported SACs were modelled by anchoring individual metal atoms onto various 2D substrates, including graphitic carbon nitride (g-C$_3$N$_4$), nitrogen-doped graphene, graphene with dual-vacancy, black phosphorous, boron nitride, and monolayer C$_2$N.
The specific single metal atoms studied are outlined in Supplementary Figure 1 (TODO:).
To prevent interactions between adjacent images along the z-axis, vacuum layers of 15 \text{\AA} were introduced.

@@

\subsection{Hybrid descriptor for volcano plots}
Hybrid descriptors incorporate both descriptors within a single equation, enhancing the accuracy of scaling relationship with minimal computational overhead, as demonstrated in Supplementary Figure 10 (TODO:) and Supplementary Table 13 (TODO:).
Comprehensive analyses are presented Supplementary Section 1.5. This leads to the formulation of the scaling relation:
\begin{equation*}
G_{\text{ads}} Z = k[\theta^* G_{\text{ads}} \ce{CO} + (1-\theta^*) G_{\text{ads}} \ce{OH}] + c_Z
\end{equation*}
Here, $\theta$ represents the mixing ratio of the CO descriptor, $\textit{Z}$ denotes any adsorbate, and $\textit{k}$ and $\textit{c}_{Z}$ are adsorbate-specific scaling parameters.
The optimal mixing ratio $\theta$ is determined through iterative exploration across the range, utilizing a step length of 0.01.

@@

\subsection{CNN architecture and hyperparameter tuning}
We designed our CNN architecture drawing inspiration from its successful implementation in tasks involving electrocardiogram signal classification \cite{weimann2021transfer}, as illustrated in \cref{main_fig2:cnn_for_eads}b and \cref{main_fig2:cnn_for_eads}c.
This multi-branched CNN comprises nine branches tailored to handle eDOS orbitals while ensuring a reduced memory footprint.

@@

\subsection{Occlusion experiments}
Supplementary Figure 21 (TODO:) visually illustrates the occlusion experiment methodology.
To maintain consistent shapes, zero-padding was applied to the input eDOS arrays.
Focusing on understanding the influence of metals on adsorption behavior, a masker with dimensions [width, 1, 1] was employed along the metal channel to eliminate cross-orbital interactions.
Here, "width" denotes the masker width, and detailed explanations for determining the width are provided in Supplementary Section 2.6.
Through recursive application of these maskers, variations in predicted adsorption energies from the CNN model were recorded, generating an occlusion array.

\subsection{Shifting experiments}
Supplementary Figure 30 (TODO:) illustrates the implemented shifting experiment protocol.
In this study, the initial electronic density of states underwent controlled shifting along the energy axis.
This protocol ranges from -1 eV to 1 eV, with a step length of 0.005 eV per image.
To maintain a uniform energy window, each image underwent cropping and padding procedures.
The resulted output arrays took the form of [400, 1] arrays for simultaneous orbital shifting and [400, 9] arrays for individual orbital shifting.
The energy range for shifting was chosen to induce a moderate disturbance to adsorption energy.
